---
title: "ROC calculation"
output: html_document
date: "2023"
---

```{r}
library(zoo)
library(tidyverse)
library(gridExtra)
```


# Load data

```{r}
#data <- read.csv("../data/nv.csv")
data <- read.csv("../data/data_week_noduplicate.csv")
```

# Predict missing values for 193 variables and viral load

Prediction is done separately for naive and vaccinated samples. (after challenge)

```{r}
# Dataframes for putting predicted values for all measurements
data.naive <- data.frame(week.timepoint = rep(seq(0,20), 5),
                  Animal.No. = unlist(lapply(as.character(unique(data[which(data$arm == 'naive'), "Animal.No."])), rep, times = 21)))
data.vac <- data.frame(week.timepoint = rep(seq(0,20), 6),
                  Animal.No. = unlist(lapply(as.character(unique(data[which(data$arm == 'SIVmac239_ΔGY' & data$challenge.tp == FALSE), "Animal.No."])), rep, times =  21)))
roc.naive <- data.frame(week.timepoint = rep(seq(0,20), 5),
                  Animal.No. = unlist(lapply(as.character(unique(data[which(data$arm == 'naive'), "Animal.No."])), rep, times = 21)))
roc.vac <- data.frame(week.timepoint = rep(seq(0,20), 6),
                  Animal.No. = unlist(lapply(as.character(unique(data[which(data$arm == 'SIVmac239_ΔGY' & data$challenge.tp == FALSE), "Animal.No."])), rep, times =  21)))
```

## Naive

```{r}
for (measurement in colnames(data)[c(5,11:203)]) {
  
# new dataframe for viral loads
viral.load.auc <- data.frame(week.timepoint = seq(0,20))
# assign week values
#viral.load.auc$week.timepoint <- seq(0,20)

# naive cases
for (i in unique(data[which(data$arm == 'naive'),]$Animal.No.)) {
  print(i)
  # subset animals for merging
  temp.df.merging <- data[which(data$Animal.No. == i),c("week.timepoint", measurement)]
  colnames(temp.df.merging) <- c("week.timepoint",i)
  
  temp.df.merging <- temp.df.merging[temp.df.merging$week.timepoint <=20,]
  
  # assign initial viral loads
  viral.load.auc <- merge(viral.load.auc, temp.df.merging, by = "week.timepoint", all = TRUE)
  #viral.load.auc[,i] <- viral.load.auc$RNACopies.ml # assign RNA values to the corresponding animal.no. column
  #viral.load.auc <- viral.load.auc[,-13] #drop RNACopies.ml column
}

# fit a curve to the data using loess (naive)
loess.pred <- data.frame(week.timepoint = seq(0,20,1))

for (i in unique(data[which(data$arm == 'naive'),]$Animal.No.)) {
  #print(i)
  #subset wrt Animal.No. and first 20 weeks
  temp.df.merging <- data[which(data$Animal.No. == i),c("week.timepoint", measurement)]
  temp.df.merging <- temp.df.merging[temp.df.merging$week.timepoint <= 20,]
  # loess fit
  fit <- loess(get(measurement) ~ week.timepoint, data = temp.df.merging)
  # prediction of missing points
  predictions.df <- data.frame(week.timepoint = seq(0,20,1))
  pred <- data.frame(vl.pred = predict(fit, predictions.df))
  pred$week.timepoint <- predictions.df$week.timepoint
  
  # assign to df
  loess.pred[,i] <- pred$vl.pred 
}
# Forecast for NA values (naive)
for (i in unique(data[which(data$arm == 'naive'),]$Animal.No.)) {
  #print(i)
  # calculate fit for forecast
  forecast.fit <- auto.arima(loess.pred[1:length(na.omit(loess.pred[,i])),i], seasonal = FALSE)
  # forecast
  forecast.number <- nrow(loess.pred) - length(na.omit(loess.pred[,i]))
  loess.pred.forecast <- forecast(forecast.fit,forecast.number)
  print(loess.pred.forecast$method)
  # append forecast to initial values
  full.set <- c(loess.pred.forecast$x, loess.pred.forecast$mean)
  # assign to df
  #print(i)
  #print(length(full.set))
  #print(full.set)
  #print("-------")
  #col.name.holder <- paste0(i, ".forecast")
  loess.pred[,i] <- full.set
}
loess.pred[loess.pred < 0] <- 0
long.holder <- pivot_longer(loess.pred, cols=-week.timepoint)
colnames(long.holder) <- c("week.timepoint", "Animal.No.", measurement)

######### ROC calculation
roc.df <- data.frame(loess.pred$week.timepoint,
                    lapply(loess.pred[,-1], function(x) c(NA, diff(x) / x[-length(x)] * 100)))
colnames(roc.df)[1] <- "week.timepoint"
roc.long.holder <- pivot_longer(roc.df, cols=-week.timepoint)
colnames(roc.long.holder) <- c("week.timepoint", "Animal.No.", paste0("roc_",measurement))
#########

#add measurement to main df (data.naive)
data.naive <- inner_join(data.naive, long.holder, by = c("week.timepoint", "Animal.No."))

#add roc to roc df (roc.naive)
roc.naive <- inner_join(roc.naive, roc.long.holder, by = c("week.timepoint", "Animal.No."))
}
roc.naive[is.na(roc.naive)] <- 0
```

### Visualize absolute measurement and ROC feature
```{r}
ggplot(data[which(data$Animal.No. == "MA39"),], aes(x= week.timepoint, y= TIgG_SIV.1.mac251)) + 
  geom_line() + 
  geom_point(data = data.naive[which(data.naive$Animal.No. == "MA39"),], aes(x= week.timepoint, y= TIgG_SIV.1.mac251))
```

```{r}
ggplot(data = roc.naive[which(roc.naive$Animal.No. == "MA39"),], aes(x= week.timepoint, y= roc_TIgG_SIV.1.mac251)) + 
  geom_point() + 
  geom_line()
```

## vaccinated (after challenge)

```{r}
for (measurement in colnames(data)[c(5,11:203)]) {
# fit a curve to the data using loess (∆GY, after challenge)
# use the same set for the loop since LJ48 does not have data before challenge
loess.pred.gy.after <- data.frame(week.timepoint = seq(0,20,1))

for (i in unique(data[which(data$arm == 'SIVmac239_ΔGY' & data$challenge.tp == FALSE),]$Animal.No.)) {
  #print(i)
  #subset wrt Animal.No. and first 20 weeks
  temp.df.merging <- data[which(data$challenge.tp == TRUE),c("week.timepoint", measurement,"Animal.No.")]
  temp.df.merging <- temp.df.merging[which(temp.df.merging$Animal.No. == i),] 
  temp.df.merging <- temp.df.merging[temp.df.merging$week.timepoint <= 20,]
  #print(temp.df.merging)
  # loess fit
  fit <- loess(get(measurement) ~ week.timepoint, data = temp.df.merging)
  # prediction of missing points
  predictions.df <- data.frame(week.timepoint = seq(0,20,1))
  pred <- data.frame(vl.pred = predict(fit, predictions.df))
  pred$week.timepoint <- predictions.df$week.timepoint
  
  # assign to df
  loess.pred.gy.after[,i] <- pred$vl.pred 
}
loess.pred.gy.after[loess.pred.gy.after <0] <- 0 #convert negative predictions to zero.

# Forecast for NA values (GY after challenge)
gy.after.challenge.cases <- c("LJ02", "LJ04", "LJ46", "LJ47", "MA18", "MA23") #LJ48 removed no measurement before challenge 

for (i in gy.after.challenge.cases) {
  print(i)
  # calculate fit for forecast
  forecast.fit <- auto.arima(loess.pred.gy.after[1:length(na.omit(loess.pred.gy.after[,i])),i], seasonal = FALSE)
  # forecast
  forecast.number <- nrow(loess.pred.gy.after) - length(na.omit(loess.pred.gy.after[,i]))
  if (forecast.number == 0) {
    next #skip this iteration (no need for forecasting for that sample)
  }
  loess.pred.forecast <- forecast(forecast.fit,forecast.number)
  print(loess.pred.forecast$method)
  # append forecast to initial values
  full.set <- c(loess.pred.forecast$x, loess.pred.forecast$mean)
  # assign to df
  #print(i)
  #print(length(full.set))
  #print(full.set)
  #print("-------")
  #col.name.holder <- paste0(i, ".forecast")
  loess.pred.gy.after[,i] <- full.set
}
loess.pred.gy.after[loess.pred.gy.after <0] <- 0 #convert negative predictions to zero.
long.holder.gy <- pivot_longer(loess.pred.gy.after, cols=-week.timepoint)
colnames(long.holder.gy) <- c("week.timepoint", "Animal.No.", measurement)

######### ROC calculation
roc.df.gy <- data.frame(loess.pred.gy.after$week.timepoint,
                    lapply(loess.pred.gy.after[,-1], function(x) c(NA, diff(x) / x[-length(x)] * 100)))
colnames(roc.df.gy)[1] <- "week.timepoint"
roc.gy.long.holder <- pivot_longer(roc.df.gy, cols=-week.timepoint)
colnames(roc.gy.long.holder) <- c("week.timepoint", "Animal.No.", paste0("roc_",measurement))
#########

#add measurement to main df (data.vac)
data.vac <- inner_join(data.vac, long.holder.gy, by = c("week.timepoint", "Animal.No."))

#add roc to roc df (roc.vac)
roc.vac <- inner_join(roc.vac, roc.gy.long.holder, by = c("week.timepoint", "Animal.No."))
}
```

```{r}
# Missing value handling
roc.vac <- do.call(data.frame, lapply(roc.vac, function(x) replace(x, is.infinite(x), NA)))
roc.vac[is.na(roc.vac)] <- 0
```

### Visualize absolute measurement and ROC feature

```{r}
ggplot(data[which(data$Animal.No. == "LJ04"),], aes(x= week.timepoint, y= TIgG_SIV.1.mac251)) + 
  geom_line() + 
  geom_point(data = data.vac[which(data.vac$Animal.No. == "LJ04"),], aes(x= week.timepoint, y= TIgG_SIV.1.mac251))
```


## vaccinated (before challenge)

```{r}
for (measurement in colnames(data)[c(5,11:203)]) {
# fit a curve to the data using loess (∆GY, after challenge)
# use the same set for the loop since LJ48 does not have data before challenge
loess.pred.gy.after <- data.frame(week.timepoint = seq(0,20,1))

for (i in unique(data[which(data$arm == 'SIVmac239_ΔGY' & data$challenge.tp == FALSE),]$Animal.No.)) {
  #print(i)
  #subset wrt Animal.No. and first 20 weeks
  temp.df.merging <- data[which(data$challenge.tp == FALSE),c("week.timepoint", measurement,"Animal.No.")] # FALSE for before challenge
  temp.df.merging <- temp.df.merging[which(temp.df.merging$Animal.No. == i),]
  if (i == "LJ47") {
    temp.df.merging <- temp.df.merging[temp.df.merging$week.timepoint <= 40,]
    temp.df.merging[3,"week.timepoint"] <- 20
  } else {
      temp.df.merging <- temp.df.merging[temp.df.merging$week.timepoint <= 20,]
  }
  #print(temp.df.merging)
  # loess fit
  fit <- loess(get(measurement) ~ week.timepoint, data = temp.df.merging)
  # prediction of missing points
  predictions.df <- data.frame(week.timepoint = seq(0,20,1))
  pred <- data.frame(vl.pred = predict(fit, predictions.df))
  pred$week.timepoint <- predictions.df$week.timepoint
  
  # assign to df
  loess.pred.gy.after[,i] <- pred$vl.pred 
}
loess.pred.gy.after[loess.pred.gy.after <0] <- 0 #convert negative predictions to zero.

# Forecast for NA values (GY after challenge)
gy.after.challenge.cases <- c("LJ02", "LJ04", "LJ46", "LJ47", "MA18", "MA23") #LJ48 removed no measurement before challenge 

for (i in gy.after.challenge.cases) {
  print(i)
  # calculate fit for forecast
  forecast.fit <- auto.arima(loess.pred.gy.after[1:length(na.omit(loess.pred.gy.after[,i])),i], seasonal = FALSE)
  # forecast
  forecast.number <- nrow(loess.pred.gy.after) - length(na.omit(loess.pred.gy.after[,i]))
  if (forecast.number == 0) {
    next #skip this iteration (no need for forecasting for that sample)
  }
  loess.pred.forecast <- forecast(forecast.fit,forecast.number)
  print(loess.pred.forecast$method)
  # append forecast to initial values
  full.set <- c(loess.pred.forecast$x, loess.pred.forecast$mean)
  # assign to df
  #print(i)
  #print(length(full.set))
  #print(full.set)
  #print("-------")
  #col.name.holder <- paste0(i, ".forecast")
  loess.pred.gy.after[,i] <- full.set
}
loess.pred.gy.after[loess.pred.gy.after <0] <- 0 #convert negative predictions to zero.
long.holder.gy <- pivot_longer(loess.pred.gy.after, cols=-week.timepoint)
colnames(long.holder.gy) <- c("week.timepoint", "Animal.No.", measurement)

######### ROC calculation
roc.df.gy <- data.frame(loess.pred.gy.after$week.timepoint,
                    lapply(loess.pred.gy.after[,-1], function(x) c(NA, diff(x) / x[-length(x)] * 100)))
colnames(roc.df.gy)[1] <- "week.timepoint"
roc.gy.long.holder <- pivot_longer(roc.df.gy, cols=-week.timepoint)
colnames(roc.gy.long.holder) <- c("week.timepoint", "Animal.No.", paste0("roc_",measurement))
#########

#add measurement to main df (data.vac)
data.vac <- inner_join(data.vac, long.holder.gy, by = c("week.timepoint", "Animal.No."))

#add roc to roc df (roc.vac)
roc.vac <- inner_join(roc.vac, roc.gy.long.holder, by = c("week.timepoint", "Animal.No."))
}
```

```{r}
# Missing value handling
roc.vac <- do.call(data.frame, lapply(roc.vac, function(x) replace(x, is.infinite(x), NA)))
roc.vac[is.na(roc.vac)] <- 0
```

### Visualize absolute measurement and ROC feature

```{r}
ggplot(data[which(data$Animal.No. == "LJ04"),], aes(x= week.timepoint, y= TIgG_SIV.1.mac251)) + 
  geom_line() + 
  geom_point(data = data.vac[which(data.vac$Animal.No. == "LJ04"),], aes(x= week.timepoint, y= TIgG_SIV.1.mac251))
```
# Dataframe for all ROC features

```{r}
# Put all predicted measurements in a single df
#roc.after.challenge <- rbind(roc.naive, roc.vac)
roc.after.challenge <- roc.vac

# add group info (vac or naive)
roc.after.challenge$arm <- data[match(roc.after.challenge$Animal.No., data$Animal.No.), "arm"]
roc.after.challenge <- roc.after.challenge[,c("Animal.No.", "arm", "week.timepoint", colnames(roc.after.challenge)[3:196])]

# Inf or NA values
roc.after.challenge <- do.call(data.frame, lapply(roc.after.challenge, function(x) replace(x, is.infinite(x), NA)))
roc.after.challenge[is.na(roc.after.challenge)] <- 0
```

```{r}
# merge with the absolute measurements
roc.after.challenge <- bind_cols(data.vac, roc.after.challenge[,c(4:197)])
```

```{r}
# save predictions
#write.csv(roc.after.challenge, "gy_before_measurement_roc_20weeks.csv", row.names = FALSE)
```

